---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r computetime}
Sys.setlocale("LC_TIME","English")
time <- format(Sys.time(), "%F %T %Z")
tz   <- Sys.timezone()
```

The current time is **`r time`** in **`r tz`**.

```{r loadlibraries}
library(knitr)
library(dplyr)
setwd('C:/Users/acer/Developer/RepData_PeerAssessment1')
```

## Loading and preprocessing the data
```{r loaddata}
if(!file.exists("activity.csv")){
        unzip("activity.zip", exdir = ".")
}
activity <- read.csv("activity.csv", header = T, sep = ',')
# Convert date from factor into a date object.
activity$datePOSIX <- as.POSIXct(activity$date, format = "%Y-%m-%d", tz = "GMT")
```

## What is mean total number of steps taken per day?
```{r meantotal}
by_day <- group_by(activity, datePOSIX)
steps <- na.omit(summarise(by_day, Total = sum(steps)))
hist(steps$Total)
mean(steps$Total, na.rm = TRUE)
median(steps$Total, na.rm = TRUE)
```


## What is the average daily activity pattern?
```{r averagedaily}

activity$datetimePOSIX <- as.POSIXct(paste(activity$date, 
                                           sprintf("%04d", activity$interval)
                                           ),
                                     format="%Y-%m-%d %H%M", tz = "GMT")

by_datetime <- group_by(activity, datetimePOSIX)
steps_mean_dt <- na.omit(summarise(by_datetime, Average = mean(steps)))

plot(steps_mean_dt, type = "l", main = "Average Daily Activity Pattern")

steps_mean_dt[ which.max(steps_mean_dt$Average), ]
```


## Imputing missing values
The total of missing values can be computed with:
```{r total NA}
totalNA <- dim(activity[ is.na(activity$steps), ])[1]
```
The value is **`r totalNA`**. The presence of missing may introduce
bias into some calculations or summaries of data.  
In order to fill these **`r totalNA`** NA values, the rounded average steps (mean)
for that 5 minute interval were taken:
```{r strategy for imputing}
IMP <- aggregate(steps ~ interval, data = activity, mean)
IMP$steps <- round(IMP$steps)
```
A data set that is equal to the original dataset but with the missing data filled in were prepared:
```{r create new data set}
names(IMP)[2] <- "activityMean"
bind <- merge(activity, IMP)
```
Perform imputing.
```{r imputing}
bind$steps[ is.na( bind$steps ) ] <- bind$activityMean[ is.na( bind$steps ) ]
bind$activityMean <- NULL
```

Data set **bind** is the data set **activity** but with the missing data filled in.
```{r data sets}
str(activity)
str(bind)
```

Make a histogram of the total number of steps taken each day and 
Calculate and report the mean and median total number of steps taken per day.

```{r report}
by_day_bind <- group_by(bind, datePOSIX)
steps_bind <- summarise(by_day_bind, Total = sum(steps))
hist(steps_bind$Total)
mean(steps_bind$Total)
median(steps_bind$Total)
```

Do these values differ from the estimates from the first part of the assignment?
What is the impact of imputing missing data on the estimates of the total daily number of steps?

## Are there differences in activity patterns between weekdays and weekends?
```{r differences}

```
