---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

```{r computetime}
Sys.setlocale("LC_TIME","English")
time <- format(Sys.time(), "%F %T %Z")
tz   <- Sys.timezone()
```

The current time is **`r time`** in **`r tz`**.

```{r loadlibraries}
library(knitr)
library(dplyr)
setwd('C:/Users/acer/Developer/RepData_PeerAssessment1')
```

## Loading and preprocessing the data
```{r loaddata}
if(!file.exists("activity.csv")){
        unzip("activity.zip", exdir = ".")
}
activity <- read.csv("activity.csv", header = T, sep = ',')
# Convert date from factor into a date object.
activity$datePOSIX <- as.POSIXct(activity$date, format = "%Y-%m-%d", tz = "GMT")
```

## What is mean total number of steps taken per day?
```{r meantotal}
by_day <- group_by(activity, datePOSIX)
steps <- na.omit(summarise(by_day, Total = sum(steps)))
hist(steps$Total)
mean(steps$Total, na.rm = TRUE)
median(steps$Total, na.rm = TRUE)
```


## What is the average daily activity pattern?
```{r averagedaily}

activity$datetimePOSIX <- as.POSIXct(paste(activity$date, 
                                           sprintf("%04d", activity$interval)
                                           ),
                                     format="%Y-%m-%d %H%M", tz = "GMT")

by_datetime <- group_by(activity, datetimePOSIX)
steps_mean_dt <- na.omit(summarise(by_datetime, Average = mean(steps)))

plot(steps_mean_dt, type = "l", main = "Average Daily Activity Pattern")

steps_mean_dt[ which.max(steps_mean_dt$Average), ]
```


## Imputing missing values
The total of missing values can be computed with:
```{r total NA}
totalNA <- sum(is.na(activity$steps))
```
The value is **`r totalNA`**. The presence of missing may introduce
bias into some calculations or summaries of data.  
In order to fill these **`r totalNA`** NA values, the rounded average steps (mean)
for that 5 minute interval were taken:
```{r strategy for imputing}
IMP <- aggregate(steps ~ interval, data = activity, mean)
IMP$steps <- round(IMP$steps)
```
A data set that is equal to the original dataset but with the missing data filled in were prepared:
```{r create new data set}
names(IMP)[2] <- "activityMean"
bind <- merge(activity, IMP)
```
Perform imputing.
```{r imputing}
bind$steps[ is.na( bind$steps ) ] <- bind$activityMean[ is.na( bind$steps ) ]
bind$activityMean <- NULL
```

Data set **bind** is the data set **activity** but with the missing data filled in.
```{r data sets}
str(activity)
str(bind)
```

Make a histogram of the total number of steps taken each day and 
Calculate and report the mean and median total number of steps taken per day.

```{r report}
by_day_bind <- group_by(bind, datePOSIX)
steps_bind <- summarise(by_day_bind, Total = sum(steps))
hist(steps_bind$Total)
mean(steps_bind$Total)
median(steps_bind$Total)
```

These values differ from the estimates from the first part of the assignment by:
```{r diff}
mean(steps$Total) - mean(steps_bind$Total)
median(steps$Total) - median(steps_bind$Total)

paste0(round((1 - (mean(steps$Total) / mean(steps_bind$Total))) * 100, digits = 2),"%")
paste0(round((1 - (median(steps$Total) / median(steps_bind$Total))) * 100, digits = 2),"%")
```
The impact of imputing missing data on the estimates of the total daily number
of steps increases precision decreasing the estimates, as shown above.

## Are there differences in activity patterns between weekdays and weekends?
For this part the weekdays() function may be of some help here. Use the dataset with the filled-in missing values for this part.

Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r new factor weekday / weekend}
bind$weekday <- as.factor(ifelse(weekdays(activity$datePOSIX) %in% c('Saturday', 'Sunday'), 'weekend', 'weekday'))
```
Make a panel plot containing a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r differences}

```
